---
title: "Developing Data Products - 2nd Week Project"
author: "Harry Ramirez"
date: "June 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

<font size="3">The objective of this project is to build a webpage that shows an interactive map developed using *R Markdown* and *leaflet*. We designed a webpage with an earthquakes map, showing details of epicenter, magnitude, date, time and location for each quake. We found current data at the [USGS](https://earthquake.usgs.gov/earthquakes/feed/v1.0/csv.php) website. Using Markdown we rendered a webpage that was hosted at [github](https://harryrampr.github.io/DDP_2ndWeekProject/webpage.html). As an extra, we added to the webpage a table rendered with *googleVis* that shows the top ranking earthquakes for the report period.

Here are the details of how the interactive map and table were created:
</font>

```{r message=FALSE}
# Packages
library(dplyr)
library(lubridate)
library(leaflet)
library(htmltools)
library(googleVis)
op <- options(gvis.plot.tag='chart')

# Report parameters
dataUrl <- "https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/1.0_month.csv"
dataFileName <- "rawData.rda"
repDays <- 15
startingLng <- -95
startingLat <- 38
startingZoom <- 2
legendLevels <- c("2.0", "3.0", "4.0", "5.0", "6.0", "7.0", "8.0+")
mapColors <- list(lightgreen = "#BBF970",
                  lightblue = "#88DAFF",
                  lightgray = "#A3A3A3",
                  pink = "#FF8EE9",
                  orange = "#F2952F",
                  red = "#D33D2A",
                  darkred = "#A03336")

# Download data if necessary
if(!file.exists(dataFileName)) {
    rawData <- read.csv(dataUrl, stringsAsFactors = F)
    save(rawData, file = "rawData.rda")
} else {
    load(file = dataFileName)
}

# Color scale earthquake magnitude
MagToCol <- function(magnitude) {
    if(magnitude < 3) {
        return(names(mapColors)[1])
    } else if(magnitude < 4) {
        return(names(mapColors)[2])
    } else if(magnitude < 5) {
        return(names(mapColors)[3])
    } else if(magnitude < 6) {
        return(names(mapColors)[4])
    } else if(magnitude < 7) {
        return(names(mapColors)[5])
    } else if(magnitude < 8) {
        return(names(mapColors)[6])
    } else {
        return(names(mapColors)[7])
    }
}

# Create icons labels
CustomLabels <- function(data) {
    magHTML <- paste0("<p style = 'margin: 0'>",
                      "<strong>Magnitude:</strong> ",
                      data$mag, "</p>")
    dateHTML <- paste0("<p style = 'margin: 0'>",
                      "<strong>Date:</strong> ",
                      data$date, "</p>")
    timeHTML <- paste0("<p style = 'margin: 0'>",
                       "<strong>Time:</strong> ",
                       data$hour, "</p>")
    locationHTML <- paste0("<p style = 'margin: 0'>",
                       "<strong>Location:</strong> ",
                       data$place, "</p>")
    paste0(magHTML, dateHTML, timeHTML, locationHTML)
}

# Select data to plot in report
quakesData <- rawData %>% 
    filter(type == "earthquake") %>% 
    filter(mag >= 2) %>%
    mutate(time = ymd_hms(time)) %>% 
    filter(time < date(max(.$time))) %>% 
    filter(time >= date(max(.$time)) - days(repDays - 1)) %>% 
    mutate(date = format(time, format = "%m/%d/%Y")) %>% 
    mutate(hour = format(time, format = "%H:%M:%S %Z")) %>% 
    mutate(label = CustomLabels(.)) %>% 
    mutate(color = lapply(mag, MagToCol)) %>% 
    select(longitude, latitude, time, date, hour, mag, place, label, color)

# Calc report dates
repStartingDate <- min(quakesData$time)
repStartingDate <- format(repStartingDate, format = "%m/%d/%Y")
repEndingDate <- max(quakesData$time)
repEndingDate <- format(repEndingDate, format = "%m/%d/%Y")
save(repStartingDate, repEndingDate, file = "repdates.rda")

# Customize map icons
mapIcons <- awesomeIcons(icon = "dot-circle-o",
                      iconColor = "black",
                      library = "fa",
                      markerColor = quakesData$color)

# Render interactive map
quakesMap <- quakesData %>% 
    leaflet() %>%
    addProviderTiles("Esri.OceanBasemap") %>%
    setView(lng = startingLng,
            lat = startingLat,
            zoom = startingZoom) %>% 
    addAwesomeMarkers(lng = ~longitude,
                      lat = ~latitude,
                      icon = mapIcons,
                      label = ~lapply(label, HTML)) %>% 
    addLegend(labels = legendLevels,
              colors = unname(unlist(mapColors)),
              opacity = 0.95)

save(quakesMap, file = "map.rda")

# Calc top 20 earthquakes
top20 <- quakesData %>% 
    arrange(desc(mag), desc(time)) %>% 
    rename(Magnitude=mag, Longitude=longitude, Latitude=latitude,
           Date=date, Time=hour, Location=place) %>% 
    mutate(Ranking=1:nrow(.)) %>% 
    select(Ranking, Magnitude, Longitude, Latitude, Date, Time, Location) %>% 
    head(20)

# Render top 20 table
top20Table <- gvisTable(top20,
                        formats=list(Magnitude="#,###.#",
                                     Longitude="#,###.##",
                                     Latitude="#,###.##"),
                        options=list(page='enable',
                                     height='automatic',
                                     width='910'))

save(top20Table, file = "table.rda")

```
<br style="margin-top: 20px">

### Earthquakes Map (`r repStartingDate` - `r repEndingDate`)

```{r out.width = '100%', echo=FALSE}
quakesMap
```
<br style="margin-top: 20px">

### Top Ranking Earthquakes (`r repStartingDate` - `r repEndingDate`)

```{r results="asis", out.width = '100%', echo=FALSE}
print(top20Table)
```
<br style="margin-top: 20px">
